Prediction of 2024 US election ...

### 2024-11-03 paper/paper.qmd
removed text
update paper, literature and renv

library(tidyverse)
library(here)
library(tibble)
library(knitr)
library(maps)
library(mapproj)
light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"
```

# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data. This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.

In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election.

@@

The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]

The remainder of this paper is structured as follows: [update this …]

@@

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 sample size(sample_size: The total number of respondents participating in the poll) \
 timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
 pollscore(A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
state(The U.S. state where the poll was conducted or focused) \
methodology(The method used to conduct the poll)\

@@

@@

@@

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned US election data 
#| echo: false
#| warning: false
cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))

cleaned_data |>
  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
  head(6) |>
  kable(
    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "Candidate Name"),
    booktabs = TRUE
    ) 
```

@@

This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix). The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary

@@

```
In figure @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump. Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.

@@

```
Figure @fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.

@@

# this model is really here just for explain, this realtionship should not
# be treated as lienar
# cor(graph_data$sample_size, graph_data$pollscore)
# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
# summary(lm_score_sample)
```
Figure @fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

@@

```{r}
#| label: fig-pollscore-canidate-pct
#| fig-cap: More Reliable Pollsters Score Higher For Trump
#| echo: false
#| warning: false

@@

```
Figure @fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage. For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support. Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise. This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.

Based on @tbl-state-candidate-support, state might have to be a removed variable.

@@

```
Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election. This is a result of non-major candidate support rapidly decreasing. Specifically, Trump support is increasing at a faster rate than Harris. However, the polls in general show a slight lead for Harris throughout the last 100 days.

@@

\newpage
# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

@@

```
The state statuses presented in this map are based on evaluations from @cnn, @foxnews, and @msnbc, which are generally agreed upon within the American political community. These organizations assessed historical voting patterns and recent polling data to derive their conclusions.

@@

The team will then find and use multiple databases, such as @uspostalserviceresidentialaddressfile, to find addresses, telephones, and emails of potential voters in those states. According to @pewresearch, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums. Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email. Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.

@@

The team's ideal survey will be made using @qualtrics. It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants. To identify participants who are not engaged, the research team has decided to add a worthless question. This question is extremely simple and clearly has one correct answer. Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data. An example of this is question number 7.

Order bias occurs when the sequence of questions subconsciously influences participants' responses. To prevent this, many questions should be randomized upon entry to the survey. However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized. The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics paid plan.

Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently. To prevent this, answer choices should be randomized upon survey entry. Questions 8 & 9 could benefit from this. Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics paid plan.

@@

Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics survey.

@@

1. Select your race(s) (racial options where chosen based on @whitehouseracialoptions-PLACEHOLDER)
   - White
   - Black or African American 
   - American Indian or Alaska Native 
   - Native Hawaiian or Pacific Islander 
   - Middle Eastern or North African
   - Asian 
   - Other 
   - Prefer not to answer

@@

3. Please enter your age (in numbers) 
  - This is a text input field. Please note that this field has the auto-validation feature set to numbers in @qualtrics. As a result, participants can only input numbers in this field and are alert if they have not.

@@

>>>>

library(tidyverse)
library(here)
library(tibble)
library(knitr)
library(arrow)
library(rstanarm)
library(brms)
library(dplyr)
library(lubridate)
library(gridExtra)
library(Matrix)
library(rstan)

light_blue <- "#ADD8E6"
dark_blue <- "#6688CC"
light_red <- "#FF7777"
dark_red <- "#B22222"
purple <- "#9370DB"

cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
```

# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, conventional aggregation models often overlook dynamic election factors, including the credibility and quality variations among poll sources.

We build a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Our model incorporates this approach, filtering polls from @fivethirtyeight based on a numeric grade that indicates each pollster's reliability.

The objective of this study is to predict the support percentages for the primary candidates, Kamala Harris and Donald Trump, using this hybrid model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability. We also quantify other polling attributes, such as sample size, poll reliability scores, and polling methodology, which serve as predictors. Our results show that this model provides a more stable and accurate forecast than traditional poll aggregation alone.

The remainder of this paper is structured as follows: @sec-data provides an overview and exploration of the data. @sec-model provides the modeling approach, including simple and multiple linear regression models and a Bayesian hierarchical model. We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.

The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], gridExtra [@gridextra], Matrix [@Matrix], Rstan [@Rstan] and lubridate [@lubridate].

## Estimand

@@

@@

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 `sample size` (sample_size: The total number of respondents participating in the poll) \
 `days_taken_from_election` (the time gap between the poll start date and the real election date i.e days_taken_from_election = real US election date - poll start date) \
 `pollscore` (A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
`state` (The U.S. state where the poll was conducted or focused) \
`methodology` (The method used to conduct the poll)\

## Data Exploration

```{r}
#| label: fig-pairs
#| fig-cap: "Pairs Plot of Key Numeric Variables Related to Percentage Support (pct)"
#| echo: false
#| warning: false
# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)

# Create the pairs plot
pairs(numeric_data)
```

In @fig-pairs, the pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships. The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity. The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering. No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.

@@

The dataset we obtained from @fivethirtyeight is accumulated from multiple polls and surveys. According to FiveThirtyEight, they aggregate polling data conducted by other firms and organizations that meets their methodological and ethical standards [@morris_trump_2024]. The dataset contains a list of questions in polls and surveys and their results. When a new poll is conducted, the poll is appended to the dataset. FiveThirtyEight assign the poll a poll_id, and each question is assigned a question_id.

In each of the polls recorded by FiveThirtyEight, all options of each question are recorded in candidate_names, while the proportion of respondents choosing that option is recorded in pct as a percentage. The pct is our primary response variable.

There are limitations in these measurements. The differences in sampling methods, wording in survey questions and systematic biases are all reflected in the outcome [@radcliffe_538s_2023]. Thus, in the dataset, FiveThirtyEight includes other variables, such as pollscore and sample_size for their prediction model in addition to the polling results for transparency and accuracy. As an example, pollscore indicates the pollster’s reliability rating, a low pollscore indicates a higher reliability rating [@silver_polls_2008].

@@

These variables combined allow researchers to reliably analyze and predict the 2024 US election result over time.

@@

The first 6 rows of the dataset is displayed in @tbl-cleaned-data.

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned US election data 
#| echo: false
#| warning: false

cleaned_data |>
  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
  head(6) |>
  kable(
    col.names = c("pct", "Sample Size", "pollscore", "Days taken from election","state", "methodology", "Candidate name"),
    booktabs = TRUE
    ) 
```

@@

This absence of polling data is not a significant concern due to the structure of the United States Electoral College (explained in detail in the Appendix). The states lacking polling data have consistently followed historical voting patterns, so predicting the winning candidate in those states is unnecessary.

@@

In @fig-avg-pct-state, historically Democratic are polling for Kamala and historically Republican states are polling for Trump. Similarly, historically swing states also appear to be close, for instance Michigan (46.9% Trump, 47.5% Harris), Nevada (47.7% Trump, 47.9% Harris), and Pennsylvania (46.9% Trump, 48.2% Harris) all are close to an even split.

@@

@fig-avg-pct-methodology shows that polls utilizing multiple communication methods to reach voters tend to have lower interquartile ranges (IQRs) in their boxplots compared to those that rely on only one or two communication methods.

@@

# this model is really here just for explain, this realtionship should not
# be treated as lienar
# cor(graph_data$sample_size, graph_data$pollscore)
# lm_score_sample <- lm(sample_size ~ pollscore, data = graph_data)
# summary(lm_score_sample)
```
@fig-pollscore-samplesize illustrates a weak positive correlation between a pollster's pollscore and the sample size of their poll. The figure also shows that most polls have a sample size around 800 to 1200 participants. It is also important to note that a few polls with exceptionally large sample sizes were excluded from the graph due to their status as clear outliers.

@@

```{r}
#| label: fig-pollscore-candidate-pct
#| fig-cap: More Reliable Pollsters Score Higher For Trump
#| echo: false
#| warning: false

@@

@fig-pollscore-candidate-pct depicts the relationship between a pollster's poll score and a candidate's percentage. For Donald Trump, a negative correlation is observed, indicating that pollsters with lower poll scores tend to assign him a higher percentage of support. Conversely, Kamala Harris shows the opposite trend: as the poll score decreases, her percentage tends to rise. This suggests that more reliable polls, characterized by lower poll scores, report higher support for Trump compared to less reliable pollsters.

@@

Based on @fig-polldate-canidate-pct, Trump and Harris are getting more votes in polls that are done closer to the election. This is a result of non-major candidate support rapidly decreasing. Specifically, Trump support is increasing at a faster rate than Harris. However, the polls in general show a slight lead for Harris throughout the last 100 days.

@@

# Model {#sec-model}

The goal of our modeling strategy is twofold. Firstly, we aim to accurately predict the support percentage (PCT) for Harris and Trump based on relevant poll data and key influencing factors. Secondly, we seek to evaluate the efficacy of different modeling approaches—from simple linear regression (SLR) to multiple linear regression (MLR) and Bayesian hierarchical models—to understand their predictive capabilities and assess the underlying relationships between variables. By comparing these models, we can determine which approach provides the most robust and reliable predictions, while considering the variability and potential uncertainty in the data.

## Model set-up {#sec-modelsetup}

The Bayesian model is implemented in R [@citeR] using the rstanarm package as described by @rstanarm. The model is run with the following specifications:

- Formula: $\text{pct} \sim \text{pollscore} + \text{days taken from election} + \text{sample size} + (1 | \text{methodology}) + (1 | \text{state})$

- Priors: Normal(0, 2.5) for all coefficients and intercept, Exponential(1) for $\sigma$

- Settings: Seed = 123, Cores = 4, Adapt delta = 0.95

@@

## Basic Model
```{r}
#| label: fig-model-pct-pollscore
#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
#| echo: false
#| warning: false

Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
just_harris_data = Harris_data |> na.omit()
lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
predictions = predict(lm_model1, just_harris_data)
# summary(lm_model1)

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()
```

We first set up a SLR model for predicting pct based on pollscore. The model is:

$\hat{\text{pct}} = \beta_0 + \beta_1 \times \text{pollscore}$

In this SLR model, the response variable is pct and the only one predictor is the pollscore. @fig-model-pct-pollscore visualizes the relationship between the actual and predicted values of percentage support (pct) for Kamala Harris, based on a simple linear regression model with pollscore as the sole predictor. Blue points represent individual comparisons between actual and predicted values. The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.

The primary concern lies in the evident dispersion of data points, which are widely spread and do not cluster closely around the line of perfect prediction (the dashed red line). This suggests that while pollscore may have some predictive capability, it does not adequately explain the variability in pct. The observed inconsistencies between actual and predicted values indicate that the relationship between pct and pollscore is likely not sufficiently captured by a linear model with just one predictor.

```{r}
#| label: fig-model-pct-pollscore-etc
#| echo: false
#| warning: false
just_harris_data = Harris_data |> na.omit()
model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
# summary(model_MLR)
just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))

predictions = predict(model_MLR, just_harris_data)
```

```{r}
#| label: fig-model-MLR
#| fig-cap: "Actual vs Predicted Percentage Support for Kamala Harris (Multiple Predictors)"
#| echo: false
#| warning: false
ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()
```

Next, we predicts pct using pollscore, days teken from election, methodology, sample size, and state as predictors. The model is:

$\hat{\text{pct}} = \beta_0 + \beta_1 \cdot \text{pollscore} + \beta_2 \cdot \text{days\_taken\_from\_election} + \beta_3 \cdot \text{methodology} + \beta_4 \cdot \text{sample\_size} + \beta_5 \cdot \text{state}$

@fig-model-MLR compares actual pct values with the predicted pct values from the MLR model. Blue points represent individual comparisons between actual and predicted values. The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.

The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct. The majority of the data points align relatively well with the red dashed line, particularly within the middle range of pct values (approximately between 40 and 60). This alignment indicates that the model performs reasonably well in this range, with predicted values correlating strongly with the actual observed outcomes.

## Bayesian Model

Next, we set up a bayesian model as the following in accordance to @sec-modelsetup:

\begin{align*}
\text{pct}_i &\sim \mathcal{N}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_1 \cdot \text{pollscore}_i + \beta_2 \cdot \text{days\_taken\_from\_election}_i + \beta_3 \cdot \text{sample\_size}_i \\
+ u_{\text{methodology}[i]} + u_{\text{state}[i]} \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta_j &\sim \text{Normal}(0, 2.5), \quad j = 1, 2, 3 \\
u_{\text{methodology}} &\sim \text{Normal}(0, \sigma_{\text{methodology}}) \\
u_{\text{state}} &\sim \text{Normal}(0, \sigma_{\text{state}})
\end{align*}

This hierarchical Bayesian model is designed to capture both the fixed effects of predictors and the random effects of grouping factors. The response variable is pct (percentage of support), predictors are pollscore, days taken from election, and sample size. A normal prior with mean 0 and standard deviation 2.5 is set for all coefficients and the intercept, scaled automatically.

```{r}
#| label: fig-bayesian-model
#| fig-cap: "Posterior Predictive Check for the Bayesian Model Predicting pct"
#| echo: false
#| warning: false
baye_model_data = just_harris_data
baye_model_data$pct = as.factor(baye_model_data$pct)
baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
baye_model_data$methodology = as.factor(baye_model_data$methodology)
baye_model_data$pct <- as.numeric(baye_model_data$pct)
# Define the Bayesian model with brms
# test and train dataset

formula <- pct ~ pollscore + days_taken_from_election + sample_size + (1 | methodology) + (1 | state)

priors = normal(0, 2.5, autoscale = TRUE)

bayesian_model_1 <- stan_glmer(
  formula = formula,
  data = just_harris_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
pp_check(bayesian_model_1)
# summary(bayesian_model_1)
```

### Posterior Predictive Checks 
@fig-bayesian-model displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.

The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides. The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data. However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.

The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data. Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy. Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.

### Train Test Validation
we implemented a train-test split validation approach to evaluate the performance of a Bayesian hierarchical model predicting percentage support (pct) for Kamala Harris. The dataset was divided into training (70%) and test (30%) sets to fit and validate the model, respectively. We used the stan_glmer function to fit the model on the training data, with predictors including pollscore, days_taken_from_election, sample_size, and random intercepts for methodology and state. After training, we generated predictions on the test set and calculated an R squared value to quantify the model's predictive accuracy. 
```{r}
#| warning: false
#| echo: false
# Split data into training and test sets
set.seed(123)
train_indices <- sample(seq_len(nrow(just_harris_data)), size = 0.7 * nrow(just_harris_data))
train_data <- just_harris_data[train_indices, ]
test_data <- just_harris_data[-train_indices, ]

# Fit the model on the training data
bayesian_model_train <- stan_glmer(
  formula = formula,
  data = train_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)

# Predict on the test data and check performance
predictions <- posterior_predict(bayesian_model_train, newdata = test_data)
```

\begin{align}
SS_{\text{total}} = \sum_{i=1}^{n} \left( y_i - \bar{y} \right)^2\\
SS_{\text{residual}} = \sum_{i=1}^{n} \left( y_i - \hat{y}_i \right)^2\\
R^2 = 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}}\\
\text{R-squared} = 0.700563202695799
\end{align}

```{r}
#| warning: false
#| echo: false
ss_total <- sum((test_data$pct - mean(test_data$pct))^2)
ss_residual <- sum((test_data$pct - colMeans(predictions))^2)
r_squared <- 1 - (ss_residual / ss_total)
# "R-squared: 0.700563202695799"
```

Where:\
\begin{itemize}
    \item \( y_i \) represents the actual values,
    \item \( \bar{y} \) is the mean of the actual values,
    \item \( \hat{y}_i \) are the predicted values,
    \item \( SS_{\text{total}} \) is the total sum of squares,
    \item \( SS_{\text{residual}} \) is the sum of squared residuals.
\end{itemize}

We could see that the R squared value equal to the 0.700563202695799 suggesting that the model captures a substantial portion of the variability in the data and demonstrates that the model performs well in explaining the variability of the pct in the test data, reflecting strong predictive capabilities.

```{r}
#| label: fig-bayesian-model-actual-predict
#| fig-cap: "Predicted vs Actual Percentage Support (pct) for Kamala Harris in the Bayesian Model"
#| echo: false
#| warning: false
plot(test_data$pct, colMeans(predictions), 
     xlab = "Actual PCT Values", 
     ylab = "Predicted  PCT Values", 
     main = "Predicted PCT vs Actual PCT")
abline(0, 1, col = "red")  # Add a 45-degree reference line
```

 By visual checking actual pct and predicted pct plot in @fig-bayesian-model-actual-predict, the points are plotted against a 45-degree line, which represents the ideal scenario where predicted values match the actual values perfectly.

```{r}
#| label: fig-bayesian-model-residual
#| fig-cap: "Residual Plot for the Bayesian Hierarchical Model Predictions"
#| echo: false
#| warning: false
residuals <- test_data$pct - colMeans(predictions)
plot(colMeans(predictions), residuals, 
     xlab = "Predicted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red")
```

In addition, the residual plot @fig-bayesian-model-residual are fairly centered around zero with no major trend, suggesting that the model is not heavily biased in its predictions.

In conculsion, the visual checks from the predicted vs. actual plot and the residual plot, there is no strong evidence that the Bayesian model is overfitting. It appears to generalize well to the data it was trained on without showing signs of capturing noise or irrelevant patterns.

@@

To predict the support percentage (PCT) for Harris and Trump, we employed a comprehensive modeling strategy involving Simple Linear Regression (SLR), Multiple Linear Regression (MLR), and a Bayesian hierarchical model. The SLR model provided a foundational analysis of the relationship between PCT and pollscore, highlighting its limited predictive power due to a lack of complexity. The MLR model improved on this by incorporating additional predictors such as days taken from the election, sample size, methodology, and state, enhancing its predictive accuracy and capturing interactions among variables. The Bayesian hierarchical model further refined our approach, incorporating prior knowledge and accounting for variability at group levels (e.g., methodology and state) through random intercepts. This model provided robust uncertainty quantification through credible intervals and demonstrated strong predictive performance with an R-squared value around 0.70, indicating it effectively explained data variability. By leveraging the strengths of these models, particularly the Bayesian approach’s interpretability and robustness, we achieved reliable predictions and a deeper understanding of the factors influencing support percentages.

# Result {#sec-result}
```{r}
#| label: fig-bayesian-model-CI
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

@fig-bayesian-model-CI displays the estimated posterior distribution for the intercept parameter in the Bayesian model. The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S. election data used in the analysis. The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.

The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range. The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.

In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S. election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%. This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.

```{r}
#| echo: false
#| warning: false

predictions2 <- posterior_predict(bayesian_model_1, newdata = just_harris_data)

predicted_means <- colMeans(predictions2)
predicted_intervals <- apply(predictions2, 2, quantile, probs = c(0.025, 0.975))

# Combine the results with the actual data
result_summary <- data.frame(
  Actual_PCT = just_harris_data$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)

ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "U.S. Election Prediction Results",
    x = "Actual PCT",
    y = "Predicted PCT"
  )

print(head(result_summary))

# Discussion {#sec-discussion}

## Summary of Findings

In this paper, we developed a hybrid election forecasting model to predict the 2024 U.S. presidential election results. Our model provides a stable and accurate forecast compared to traditional methods by combining poll-of-polls aggregation with additional factors such as poll score, sample size, days from the election, state, and methodology.

Our approach in models applies simple linear regression, multiple linear regression and Bayesian hierarchical modeling that includes both fixed and random effects, capturing variations in poll reliability. The results shows that a model accounting for poll quality, timing, and methodology can produce forecasts that are more resilient to biases found in individual polls.

## Insights on Polling and Election Dynamics

One of the main takeaways from this study is the impact of poll quality on election forecasts. We weight and filtered the polls by their reliability. This approach highlights the importance of screening and adjusting for poll reliability in election forecasts, particularly in a high-stakes environment where even minor polling biases can influence public perception and campaign strategies.

## Limitations of the Model

While this model presents a more comprehensive approach, there are limitations. First, Any systematic biases inherent in those polls can carry over into the model’s predictions because we are using existing poll data. For instance, if certain demographic groups are consistently underrepresented, our model may not fully capture their impact on election outcomes.

Second, the model only reflect current voting dynamics as it is based on historical data. This model does not account for short-term variability such as Biden's exit. This limitation is important in the final weeks before an election when small shifts in polling data can produce exaggerated effects.

## Future Directions
To enhance the reliability and robustness of this forecasting model, future work should focus on validating its performance across multiple election cycles. We includes previous year's polling dataset in this R project. By applying the model to past elections, researchers can assess its accuracy in different political contexts and electoral dynamics. This approach would show any limitations specific to certain election conditions, such as shifts in voter demographics or the influence of emerging media platforms on public opinion. Expanding the model’s validation across multiple election cycles could help verify its robustness and adaptability, ultimately refining its accuracy and reliability for future applications in election forecasting.

@@

@@

The state statuses presented in this map are based on evaluations from @cnn_electoral_nodate, @fox_news_2024_2024, and @nbc_news_choose_nodate, which are generally agreed upon within the American political community. These organizations assessed historical voting patterns and recent polling data to derive their conclusions.

@@

The team will then find and use multiple databases, such as @usps_customer_nodate, to find addresses, telephones, and emails of potential voters in those states. According to @pew_research_center_us_nodate, gathering this information reduces non-response and selection bias, as the team can contact the same individual through multiple mediums. Moreover, certain demographics may prefer specific communication forms; for instance, the elderly may prefer phone or paper mail polls over text or email. Additionally, to encourage individuals to complete the survey, one in every 50 participants will have a chance at winning $20.

@@

The team's ideal survey will be made using @qualtrics_qualtrics_nodate. It will attempt to ensure four things: no leading questions, no question order bias, no answer order bias, and the survey should identify non-engaged participants. To identify participants who are not engaged, the research team has decided to add a worthless question. This question is extremely simple and clearly has one correct answer. Therefore, if a participant selects the wrong answer, they most likely are not engaged with the survey and their responses should be removed from the final data. An example of this is question number 7.

Order bias occurs when the sequence of questions subconsciously influences participants' responses. To prevent this, many questions should be randomized upon entry to the survey. However, some questions must follow a specific order, such as question 11, while others like questions 1-4 can be randomized. The team's hypothetical survey has not implemented this feature, though it is available with the @qualtrics_qualtrics_nodate paid plan.

Answer bias, like order bias, occurs when the order of answer choices influences participants' selections, with the first option often chosen more frequently. To prevent this, answer choices should be randomized upon survey entry. Questions 8 & 9 could benefit from this. Likewise, the team's hypothetical survey has not implemented this feature, but it is available in the @qualtrics_qualtrics_nodate paid plan.

@@

Click this [link](https://qualtricsxm7d2hxss4j.qualtrics.com/jfe/form/SV_1Tu3PT2eUEa1Op8) for the @qualtrics_qualtrics_nodate survey.

@@

1. Select your race(s) (racial options where chosen based on @orvis_omb_2024)
   - White
   - Black or African American 
   - American Indian or Alaska Native 
   - Native Hawaiian or Pacific Islander 
   - Middle Eastern or North African
   - Asian 
   - Other 
   - Prefer not to answer

@@

3. Please enter your age (in numbers) 
  - This is a text input field. Please note that this field has the auto-validation feature set to numbers in @qualtrics_qualtrics_nodate. As a result, participants can only input numbers in this field and are alert if they have not.

@@

# Methodology of YouGov
YouGov's methodology documentations are seperated in two articles. The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.
